# tidyverse essentials (dplyr & tidyr)



## Carrega alguns pacotes a serem usados

```{r}
# Install packages and load package
#install.packages("tidyverse") # tidyverse packages
#install.packages("dplyr")     # individual package installation
#install.packages("tidyr")     # individual package installation
#install.packages("ggplot2")   # we need it for the data set
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Descrição dos dados `mpg` 


Fuel economy data from 1999 to 2008 for 38 popular models of cars
Description This dataset contains a subset of the fuel economy data that the EPA makes available on https://fueleconomy.gov/. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car. A data frame with 234 rows and 11 variables:

  - manufacturer manufacturer name

  - model model name

  - displ engine displacement, in litres

  - year year of manufacture

  - cyl number of cylinders

  - trans type of transmission

  - drv the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

  - cty city miles per gallon

  - hwy highway miles per gallon

  - fl fuel type

  - class "type" of car

```{r}
#help("mpg")
df <- mpg
#View(df) # check table
glimpse(df)#YO
```

## select()

```{r}
#   Extract columns car: manufacturer, model, year
select(df, manufacturer, model, year) # just extract columns
df.car.info <- select(df, manufacturer, model, year)
df.car.info
rm(df.car.info)

#   Extract columns that begin with letter "m"
select(df, starts_with("m"))

#   Extract columns that contain letter "r"
select(df, contains("r"))

#   Extract columns that end with letter "y"
select(df, ends_with("y"))

#   Select column by columns index
select(df, 1:3)      # first three columns
select(df, c(2,5,7)) # second, fifth and seventh column
select(df, (ncol(df)-2):ncol(df)) # last three columns

select(df,manufacturer:cyl) #YO

select(df,-(manufacturer:cyl)) #YO
``` 


##   rename()

```{r}
df1 <- rename(df, 
              mnfc = manufacturer,
              mod = model)


# select columns and rename in one take
select(df,
       mnfc = manufacturer,
       mod = model,
       everything()) # selects all other columns
```

## YO hice todo lo que está abajo!

```{r}
  df %>% 
    filter(between(cty,15, 22)) %>%
    select(cty) #YO

select(df, matches("[abc]")) #YO
select(df, starts_with("m"), starts_with("c")) #YO
select(df, ends_with("l"), ends_with("s")) #YO
```

