# tidyverse essentials (dplyr & tidyr)


Para considerar

  > Preview HTML	Ctrl+Shift+K	Command+Shift+K (R Markdown)

  > %>% is Ctrl + Shift + M (Windows) or Cmd + Shift + M (Mac).

  > You can quickly insert chunks into your R Markdown file with the keyboard shortcut Cmd + Option + I (Windows Ctrl + Alt + I).

```{r}
rm(list = ls())
graphics.off()

# Install packages and load package
#install.packages("tidyverse") # tidyverse packages
#install.packages("dplyr")     # individual package installation
#install.packages("tidyr")     # individual package installation
#install.packages("ggplot2")   # we need it for the data set

library(dplyr)
library(tidyr)
library(ggplot2)
```

## Inspect data
```{r}
#help("mpg")

df <- mpg
#View(df) # check table
print(df) # print to console
str(df) # table structure
nrow(df); ncol(df) # rows ~ cols
glimpse(df)#YO
names(df)#YO
```

## select() - Columns selection

```{r}
#   Extract columns car: manufacturer, model, year
select(df, manufacturer, model, year) # just extract columns
df.car.info <- select(df, manufacturer, model, year)
rm(df.car.info)

#   Extract columns that begin with letter "m"
select(df, starts_with("m"))

#   Extract columns that contain letter "r"
select(df, contains("r"))

#   Extract columns that end with letter "y"
select(df, ends_with("y"))

#   Select column by columns index
select(df, 1:3)      # first three columns
select(df, c(2,5,7)) # second, fifth and seventh column
select(df, (ncol(df)-2):ncol(df)) # last three columns

select(df,manufacturer:cyl) #YO

select(df,-(manufacturer:cyl)) #YO
# rename() - Rename columns
``` 


##   Rename columns "manufacturer" and "model"

```{r}
df1 <- rename(df, 
              mnfc = manufacturer,
              mod = model)


# select columns and rename in one take
select(df,
       mnfc = manufacturer,
       mod = model,
       everything()) # selects all other columns
```

## YO hice todo lo que está abajo!

```{r}
  df %>% 
    filter(between(cty,15, 22)) %>%
    select(cty) #YO

select(df, matches("[abc]")) #YO
select(df, starts_with("m"), starts_with("c")) #YO
select(df, ends_with("l"), ends_with("s")) #YO
```

